### 2025-03-02 contents/methods.tex

\section{Methods}
\label{sec:methods}

% NOT finished

to write