### 2025-08-20 contents/abstract.tex
update official phase abstract

\begin{abstract}

% NOT finished

Aim: We introduce a reliability‑aware hierarchical learning framework for automated electrocardiogram(ECG)-based Chagas cardiomyopathy screening developed by Team Revenger for the George B. Moody PhysioNet Challenge 2025, focusing on maximizing positive case retrieval under low predicted‑prevalence constraints.

Methods: to write

Results: to write

Conclusion: to write

\end{abstract}