### 2025-08-20 contents/methods.tex

\section{Methods}
\label{sec:methods}

% NOT finished

to write

\cite{wagner2020ptb_xl}

\cite{ribeiro2020automatic}

\cite{Cardoso_2016_Sami-Trop}

\cite{torch_ecg_paper}