### 2025-08-22 contents/introduction.tex

\section{Introduction}
\label{sec:intro}

% almost finished

Addressing underdiagnosis of Chagas disease through scalable ECG-based screening is the focus of the 2025 George B. Moody PhysioNet Challenge \cite{goldberger2000physionet, cinc2025}. Enabled by aggregated multi-cohort ECG datasets \cite{ribeiro2020automatic, Cardoso_2016_Sami-Trop, wagner2020ptb_xl, Nunes_2021_Incidence, Filho_2020_Prognostic}, the Challenge frames a multi-source learning setting with heterogeneous label reliability and severe class imbalance.

In this work, we propose a reliability-aware hierarchical framework that prioritizes expert-confirmed labels and mitigates severe class imbalance within a deep ECG model, with optimization aligned to prevalence-constrained sensitivity objectives.