### 2025-08-24 contents/methods.tex
finish methods section

\section{Methods}
\label{sec:methods}

% almost finished

\subsection{Datasets and Preprocessing}
\label{subsec:data}

% almost finished

We used three ECG datasets for model training, with substantial differences in sample size, chagas prevalence, and label provenance as summarized in Table~\ref{tab:dataset_stats}.

\begin{table}[!htp]
\centering
\begin{tabular}{rlll}
\toprule
Dataset & \multicolumn{1}{c}{Size} & \multicolumn{1}{c}{Chagas rate} & Label provenance \\
\midrule
SaMi-Trop & 1\,631   & 100 & expert-confirmed \\
CODE-15\% & 345\,779 & 1.795   & self-reported \\
PTB-XL    & 21\,799  & 0   & N/A \\
\bottomrule
\end{tabular}
\caption{Dataset statistics and label provenance. Chagas rate is the proportion of recordings labeled positive in each dataset. N/A indicates that confirmed Chagas cases are not expected (non-endemic population).}
\label{tab:dataset_stats}
\end{table}

All ECGs were uniformly resampled to 400 Hz, bandpass filtered (0.5–45 Hz), and z-score normalized to zero mean and unit variance computed as in Eq.~\ref{eq:z_score}:
\begin{equation}
\label{eq:z_score}
\tilde{\mathbf{x}} = \frac{\mathbf{x} - \boldsymbol{\mu}_x}{\boldsymbol{\sigma}_x}
\end{equation}
where $\mathbf{x}$ is the original ECG signal, $\boldsymbol{\mu}_x$ and $\boldsymbol{\sigma}_x$ are the mean and standard deviation of $\mathbf{x}$, respectively. We excluded ECGs shorter than 1,200 samples to ensure inputs contain enough cardiac cycles for stable model analysis.

\subsection{Reliability-Aware Hierarchical Supervision}
\label{subsec:reliability}

% almost finished

We introduce a hierarchical supervision scheme that encodes source reliability through stratified label smoothing and adaptive upsampling. Three reliability levels are defined: (1) expert-confirmed positives (SaMi-Trop, maximal trust), (2) self-reported samples (CODE-15\%, both positives and negatives, higher uncertainty), and (3) non-endemic negatives (PTB-XL, very low true prevalence but still mildly regularized). Given a one-hot label $\mathbf{y}$ ($[0, 1]$ for positives, $[1, 0]$ for negatives) and number of classes $C = 2,$ the smoothed target is computed as in Eq.~\ref{eq:label_smoothing}:
\begin{equation}
\label{eq:label_smoothing}
\tilde{\mathbf{y}} = (1 - \varepsilon) \cdot \mathbf{y} + \frac{\varepsilon}{C} \cdot \mathbf{1},
\end{equation}
where $\varepsilon$ is the smoothing factor which depends on the reliability level: 0.0 (SaMi positives), 0.6 (CODE-15\% positives \& negatives), 0.2 (PTB-XL negatives). This attenuates overconfident gradients for noisier or potentially misreported labels while preserving sharp supervision on expert-confirmed cases.

Severe class imbalance was mitigated by upsampling positives during training: positive samples from CODE-15\% were upsampled by a factor of 3, and those from SaMi-Trop by 7. No upsampling was applied to PTB-XL, which contains no positives. Smoothed labels and upsampling strategies for each dataset are summarized in Table~\ref{tab:augmentation}.

\begin{table}[!htp]
\centering
\begin{tabular}{lcccc}
\toprule
\multirow{2}{*}{Dataset} & Upsampling & \multicolumn{2}{c}{Smoothed labels} \\ \cmidrule(lr){3-4}
& factor & negative & positive \\
\midrule
SaMi-Trop & $7$ & N/A           & $[0,\,1]$     \\
CODE-15\% & $3$ & $[0.7,\,0.3]$ & $[0.3,\,0.7]$ \\
PTB-XL    & $1$ & $[0.9,\,0.1]$ & N/A           \\
\bottomrule
\end{tabular}
\caption{Smoothed labels (computed from Eq.~\ref{eq:label_smoothing}) and upsampling strategies for each dataset.}
\label{tab:augmentation}
\end{table}

\subsection{Model Architecture}
\label{subsec:model}

% almost finished

We build upon the 1D ResNet ECG classifier of Ribeiro et al.~\cite{ribeiro2020automatic} and introduce three modifications.

\textbf{(1) Bottleneck residual blocks.} We replace basic ResNet blocks with bottleneck blocks of the form $1$–$15$–$1$ (pointwise--temporal--pointwise). The middle temporal convolution (kernel size 15) applies stride 4 for downsampling; the two 1$\times$1 convolutions first reduce then expand channels with an expansion factor of 4. A projection 1$\times$1 convolution (stride 4) is used in the residual branch whenever temporal resolution or channel width changes. Across the four blocks, the reduced (bottleneck) channel widths are $128 \to 192 \to 256 \to 320$,yielding expanded output widths $512 \to 768 \to 1024 \to 1280$.

\textbf{(2) Global squeeze-and-excitation (SE).} After the final bottleneck block, a single global SE module (reduction ratio 8) \cite{hu2018senet} performs temporal average pooling to a channel descriptor, applies a two-layer bottleneck multi-layer perceptron (MLP) $1280 \to 160 \to 1280$ with ReLU and sigmoid gating, and rescales the feature map channel-wise.

\textbf{(3) Global pooling head for variable input length.} Instead of flattening a fixed-length feature map as in the original baseline, we apply global max pooling over the remaining temporal dimension, yielding a 1280-dimensional vector irrespective of input length $L$. This vector is fed to a lightweight two-layer classification MLP: a hidden fully connected layer ($1280 \to 1024$) with non-linear activation and dropout (rate 0.2), followed by a final linear layer ($1024 \to 2$) producing class logits.

\textbf{Stem and regularization.} A stem Conv1d (kernel size 15, stride 1, 64 channels) with BatchNorm and ReLU precedes the bottleneck stack. Within each bottleneck block we apply BatchNorm+ReLU after the first two convolutions and dropout (rate 0.2) after each of those activations. All convolutions use “same” padding to preserve temporal length before downsampling operations.

Overall, the changes primarily enable variable-length inference and add global channel reweighting with minimal structural overhead. The resulting model architecture is illustrated in Fig.~\ref{fig:model}.

\begin{figure*}
\centering
\input{tikz_plots/resnet_nature_comm_bottle_neck}
\caption{Model architecture. Left: overall network: a stem (Conv1d, kernel size (ks) 15, 64 channels (Ch), Batch Normalization (BN), ReLU) followed by four bottleneck residual blocks, a global squeeze-and-excitation (SE) module, global max pooling, and an MLP head producing $C = 2$ logits. Channel widths shown ($512 \to 768 \to 1024 \to 1280$) are the expanded (post-bottleneck) channels.
Right: internal bottleneck structure (1--15--1 pointwise--temporal--pointwise). The middle 15-length convolution uses stride 4 for temporal downsampling; 1$\times$1 convolutions reduce then expand channels, and a projection 1$\times$1 (stride 4) aligns resolution and width for the residual path. For clarity, dropout layers present in the implementation are omitted. Abbreviations: ks kernel size; Ch channels; BN Batch Normalization; SE squeeze-and-excitation; MLP multi-layer perceptron.}
\label{fig:model}
\end{figure*}

\subsection{Training and Implementation Setup}
\label{subsec:train}

% almost finished

We employed an asymmetric loss (ASL) \cite{ridnik2021asymmetric_loss} to further complement the reliability-aware label smoothing strategy, jointly addressing the challenges of severe class imbalance. Let $z$ denote the logits and $p=\sigma(z)$ the predicted probability of the Chagas-positive class, the ASL is defined in Eq.~\ref{eq:asl} with separate focusing parameters for positives and negatives and a clipped negative probability:
\begin{equation}
\label{eq:asl}
\begin{multlined}
L = -y \cdot (1-p)^{\gamma_{+}} \log(p) \\
\phantom{L = } - (1-y) \cdot (p_m)^{\gamma_{-}} \log(1-p_m),
\end{multlined}
\end{equation}
where $p_m = \max(p - m, 0),$ $(\gamma_{+},\gamma_{-})=(1,4)$ and margin $\delta=0.05$. We train for 30 epochs with batch size 128 using AdamW (AMSGrad, initial learning rate $1\times10^{-4}$, peak $6\times10^{-4}$ under a One-Cycle schedule, weight decay $1\times10^{-2}$). Early stopping (patience 10 epochs, monitored on a fixed 20\% internal hold-out subset) selects the final model via the Challenge metric. Each training segment is a uniform random crop (or center trim if shorter) of length 4,096 samples. All components (data pipeline, optimization utilities) are implemented using the \texttt{torch-ECG} framework \cite{torch_ecg_paper}.