### 2025-08-24 tikz_plots/resnet_nature_comm_bottle_neck.tex

\begin{tikzpicture}[
    % font=\small,
    line width=0.75pt,
    >={Latex[length=3mm]},
    node distance=6.4mm,
    every node/.style={inner sep=2pt},
    stem/.style={draw, rounded corners=2pt, minimum width=30mm, minimum height=10mm,
                 align=center, fill=blue!10},
    block/.style={draw, rounded corners=2pt, minimum width=30mm, minimum height=12mm,
                  align=center, fill=orange!12},
    se/.style={draw, rounded corners=2pt, minimum width=30mm, minimum height=10mm,
               align=center, fill=green!12},
    pool/.style={draw, rounded corners=2pt, minimum width=30mm, minimum height=10mm,
                 align=center, fill=cyan!10},
    head/.style={draw, rounded corners=2pt, minimum width=30mm, minimum height=10mm,
                 align=center, fill=purple!10},
    output/.style={draw, rounded corners=2pt, minimum width=24mm, minimum height=10mm,
                align=center, fill=red!10},
    arrow/.style={-{Latex[length=3mm]}, thick},
    darr/.style={-{Latex[length=2.4mm]}, thick},
    dashedarrow/.style={-{Latex[length=2.6mm]}, thick, dashed},
    ann/.style={inner sep=1pt, align=center},
    layer/.style={draw, rounded corners=2pt, minimum width=18mm, minimum height=6.2mm,
                  align=center, fill=blue!6},
    bnact/.style={draw, rounded corners=2pt, minimum width=18mm, minimum height=6.2mm,
                  align=center, fill=violet!10},
    act/.style={draw, rounded corners=2pt, minimum width=18mm, minimum height=6.2mm,
                  align=center, fill=red!10},
    drop/.style={draw, rounded corners=2pt, minimum width=18mm, minimum height=6.2mm,
                  align=center, fill=gray!15, dashed},
    sum/.style={draw, circle, inner sep=1.6pt, fill=white},
    % brace/.style={decorate, decoration={brace, amplitude=6pt}},
    detailbox/.style={draw, rounded corners=3pt, inner sep=6pt, fill=orange!3}
]

% ================= Left (overall vertical path) =================
\node[stem] (input) {Input \\ (12$\times$L)};
\node[stem, below=of input] (stem) {Stem \\ Conv ks=15 \\ BN + ReLU \\ 64 Ch};

% Compressed single bottleneck block (represents 4 stacked blocks)
\node[block, below=of stem] (bottleneck) {$4 \times$ Bottleneck Block \\ ks: 1--15--1 (stride=4) \\ Ch: $512 \to 768 \to 1024 \to 1280$};

% Global SE
\node[se, below=of bottleneck] (gse) {Global SE \\ reduction=8};

% Global pooling
\node[pool, below=of gse] (gap) {Global MaxPool};

% MLP Head
\node[head, below=of gap] (mlp) {MLP Head \\ (1280 $\rightarrow$ C=2)};

% Output
\node[output, below=of mlp] (out) {Output Logits};

% Arrows (overall path)
\draw[arrow] (input) -- (stem);
\draw[arrow] (stem) -- (bottleneck);
\draw[arrow] (bottleneck) -- (gse);
\draw[arrow] (gse) -- (gap);
\draw[arrow] (gap) -- (mlp);
\draw[arrow] (mlp) -- (out);

% ================= Right (expanded single bottleneck internals) =================
% Anchor for expansion (to the right of overall bottleneck block)
\coordinate (expandAnchor) at ($(input.east)+(45mm,0)$);

\node[layer, below=0mm of expandAnchor] (c1) {Conv 1$\times$1};
\node[below right=-1mm and 8mm of c1.south] {reduce};
\node[bnact, below=5mm of c1] (bn1) {BN};
\node[act, below=5mm of bn1] (a1) {ReLU};

\node[layer, below=5mm of a1] (c2) {Conv 15\\(stride 4)};
\node[bnact, below=5mm of c2] (bn2) {BN};
\node[act, below=5mm of bn2] (a2) {ReLU};
% \node[drop, below=5mm of a2] (drp) {Dropout 0.2};

\node[layer, below=5mm of a2] (c3) {Conv 1$\times$1};
\node[below right=-1mm and 8mm of c3.south] {expand};
\node[bnact, below=5mm of c3] (bn3) {BN};

\node[sum, below=5mm of bn3] (add) {+};
\node[act, below=5mm of add] (outact) {ReLU (output)};

% Projection/skip branch (since stride=4 or channel change)
\node[layer, right=10mm of c2, yshift=-7mm, minimum width=18mm] (proj) {Proj 1$\times$1 \\ (stride 4)};
\node[bnact, below=5mm of proj, minimum width=18mm] (projbn) {BN};

% Connections internal
\draw[darr] (c1) -- (bn1);
\draw[darr] (bn1) -- (a1);
\draw[darr] (a1) -- (c2);
\draw[darr] (c2) -- (bn2);
\draw[darr] (bn2) -- (a2);
% \draw[darr] (a2) -- (drp);
\draw[darr] (a2) -- (c3);
\draw[darr] (c3) -- (bn3);
\draw[darr] (bn3) -- (add);
\draw[darr] (add) -- (outact);

% Skip path
\draw[darr] (c1.east) -| ($(proj.north)+(0,0mm)$);
\draw[darr] (proj) -- (projbn);
\draw[darr] (projbn.south) |- (add.east);

\begin{scope}[on background layer]
\node[detailbox, fit=(c1) (bn1) (a1) (c2) (bn2) (a2) (c3) (bn3) (proj) (projbn) (add) (outact)] (detail) {};
\end{scope}

\draw[dashed, line width=0.9pt] (bottleneck.north east) -- (detail.north west);
\draw[dashed, line width=0.9pt] (bottleneck.south east) -- (detail.south west);

\end{tikzpicture}