### 2025-08-24 contents/results.tex
finish results section

\section{Results}
\label{sec:results}

% almost finished

The Challenge score of our team ``Revenger'' on the hidden validation set was 0.245, ranking 170th among 350 submissions. The score on the internal hold-out of the public training data, the hidden validation score, and the validation ranking are summarized in Table~\ref{tab:scores}.

\begin{table}[!htp]
\centering
\begin{tabular}{r|r|r|r}
Training & Validation & Test & Ranking \\\hline
$0.451 \pm 0.005$ & 0.245 & TBA & 170/350 \\\hline
\end{tabular}
\caption{Challenge scores for our submitted entries (team ``Revenger''). Training: internal hold-out mean $\pm$ std over repeated runs. Validation: best among 10 validation submissions. Test: to be announced. Ranking: position on the hidden validation leaderboard.}
\label{tab:scores}
\end{table}