### 2025-10-08 contents/results.tex

\section{Results}
\label{sec:results}

% almost finished

The Challenge score of our team ``Revenger'' on the hidden test set was 0.163, ranking 32nd among 40 eligible teams. This score and ranking, along with extra scores on the internal hold-out of the public training data, and on the hidden validation set, are summarized in Table~\ref{tab:scores}.

\begin{table}[!htp]
\centering
% \input{tables/official_phase_results}
\input{tables/final_results}
\caption{Challenge scores for our submitted entries (team ``Revenger''). Training: internal hold-out mean $\pm$ std over repeated runs. Validation: best among 10 validation submissions. Test: the unique test submission. Ranking: position on the hidden test leaderboard.}
\label{tab:scores}
\end{table}