### 2025-10-31 preamble.tex
update recent review opinion and make corr. changes.

\usepackage[T1]{fontenc}
\usepackage{mathtools, graphicx, xurl, hyperref}
% \usepackage{tikz-cd}
\usepackage{tikz}
\usetikzlibrary{shapes, arrows.meta, decorations.pathmorphing, decorations.pathreplacing, backgrounds, positioning, fit, calc, shadows, shapes.misc}
% \tikzexternalize[prefix=tikz/,optimize command away=\includepdf]
\usepackage{boldline, multirow}
\usepackage{booktabs} % Allows the use of \toprule, \midrule and \bottomrule in tables
\usepackage{colortbl}
\usepackage{subcaption}
\usepackage{relsize}
\usepackage{fontawesome5}
\usepackage[ruled, vlined]{algorithm2e}

\usepackage{tabularx} % for 'X' col. type and 'tabularx' environment
\usepackage{ragged2e} % for '\RaggedRight' macro
\newcolumntype{L}{>{\RaggedRight}X} % disable full justification

\usepackage{adjustbox}

% \tikzstyle{smallcircleblock} = [circle, draw, text width = 0.7em, text centered]

% \tikzstyle{wideblock} = [rectangle, draw, text width = 6.5em, text centered, rounded corners, inner sep = 7pt, minimum height = 1.0em]

\newcommand\ImageNode[4][]{
  \node[#1] (#2) {\includegraphics[#3]{#4}};
}
\definecolor{arrowblue}{RGB}{98,145,224}

\newcolumntype{g}{>{\columncolor[gray]{0.8}}l}

\newif\ifcoloredtext
\coloredtextfalse

\newif\ifboxednn
\boxednntrue

% --- unify spacing between floats and text (single-/double-column) ---
\setlength{\textfloatsep}{6pt plus 2pt minus 2pt}   % text <-> single-column floats (table/figure)
\setlength{\intextsep}{6pt plus 2pt minus 2pt}      % inline [h] floats
\setlength{\floatsep}{6pt plus 2pt minus 2pt}       % float <-> float (single-column)

\setlength{\dbltextfloatsep}{6pt plus 2pt minus 2pt} % text <-> double-column floats (figure*)
\setlength{\dblfloatsep}{6pt plus 2pt minus 2pt}     % float <-> float (double-column)

% --- caption spacing for tables only ---
\captionsetup[table]{aboveskip=6pt, belowskip=0pt}

\newenvironment{indentedquote}[1]%
{\list{}{\leftmargin=#1\rightmargin=#1}\item[]}%
{\endlist}

\newcommand\wordcount{\input{|"texcount -inc -sum -0 -utf8 -ch -template={SUM} \currfilepath"}}